Article data analysis 2

### 2019-02-07 notebooks/article_notebook2.Rmd
minor changes

@@

## Chi Squared test for significant differences in resistance mechanisms
```{r}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    unlist() %>% unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

do_chisq_test <- function(df) {
  df <- df %>%
    dplyr::slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
          dplyr::filter(Var2 < Var1) %>% t) %>%
    mutate(test = rep(1:(n()/2), each = 2)) %>%
    group_by(test) %>%
    dplyr::do(data_frame(gene = .$key,
                  test = dplyr::first(.$test),
                  species1 = dplyr::first(.$species),
                  species2 = dplyr::last(.$species),
                  data = list(matrix(c(.$Absent,
                                       .$Present),
                                     ncol = 2)))) %>%
    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
    mutate(p.value = map_dbl(chi_test, "p.value"),
           dupl = duplicated(test)) %>%
    dplyr::filter(dupl == FALSE) %>%
    ungroup() %>%
    select(gene, species1, species2, p.value) %>%
    mutate(p.value = round(p.value, 5)) %>%
    dplyr::filter(p.value <= 0.05)
  return(df)
}

@@

@@

@@

## Correlation plot for resistance mechanisms
```{r, fig.height=8, fig.width=7, warning=FALSE}
create_corr_matrix <- function(df) {
  df <- df %>%
    as.matrix %>%
    cor %>%
    as.data.frame %>%
    rownames_to_column(var = "var1") %>%
    gather(var2, value, -var1) %>%
    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
           var2 = gsub("(.*?)_R$", "\\1", var2))
  return(df)
}

total_df <- acquired_report %>%
  left_join(mut_report, by = "id") %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  select(-nt_change) %>%
  mutate_at(vars(-c(id,species)),
            funs(as.numeric(.))) %>%
  mutate(qnr = ifelse(rowSums(.[,3:9]) > 0, 1, 0)) %>%
  select(-c(id,qnrA1,qnrB19,qnrB6,qnrB60,qnrS1,qnrS2,qnrS4)) %>%
  split(., f = .$species) %>%
  lapply(., function(x) select(x, -species))

corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species") %>%
  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
           gene2 = gsub("(.*?)_.+", "\\1", var2),
           var1 = gsub(".+_(.*?)", "\\1", var1),
           var2 = gsub(".+_(.*?)", "\\1", var2),
           type1 = case_when(gene1 == "gyrA" ~ 1,
                          gene1 == "gyrB" ~ 2,
                          gene1 == "parC" ~ 3,
                          gene1 == "parE" ~ 4,
                          gene1 == "marR" ~ 5,
                          gene1 == "marA" ~ 6,
                          gene1 == "rpoB" ~ 7,
                          gene1 == "soxS" ~ 8,
                          gene1 == "robA" ~ 9,
                          gene1 == "gadX" ~ 10,
                          gene1 == "qnr" ~ 11,
                          gene1 == "qepA4" ~ 12),
           type2 = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "marA" ~ 6,
                          gene2 == "rpoB" ~ 7,
                          gene2 == "soxS" ~ 8,
                          gene2 == "robA" ~ 9,
                          gene2 == "gadX" ~ 10,
                          gene2 == "qnr" ~ 11,
                          gene2 == "qepA4" ~ 12))

@@

ggplot(corr_matrices, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradientn(colors=cols,
                       na.value="grey95",
                       limits=c(-1, 1),
                       guide=guide_colourbar(ticks=T,nbin=10,
                                             barheight=.5,label=T, 
                                             barwidth=10))+
  labs(fill = NULL)+
  theme_classic()+
  theme(axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
        axis.title = element_blank(),
        legend.position="bottom",
        legend.justification="center",
        legend.direction="horizontal")+
  coord_fixed()+
  facet_wrap(~species)
```

@@

## cgMLST-trees
```{r, fig.width=28, fig.height=28, message = FALSE, warning = FALSE}

@@

cgMLSTdata <- read.table("../data/cgMLST.txt",
                         sep = "\t",
                         row.names=1,
                         colClasses = "factor",
                         header = T)

@@

data_broiler <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% broiler_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

data_pig <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% pig_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

data_fox <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% fox_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

data_bird <- cgMLSTdata %>%
  rownames_to_column("id") %>%
  filter(id %in% bird_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id"))

@@

cgMLST_tree_broiler <- as.phylo(hclust(daisy(cgMLSTdata_broiler,
                                             metric="gower")))

@@

cgMLST_tree_pig <- as.phylo(hclust(daisy(cgMLSTdata_pig,
                                         metric="gower")))

@@

cgMLST_tree_fox <- as.phylo(hclust(daisy(cgMLSTdata_fox,
                                         metric="gower")))

@@

cgMLST_tree_bird <- as.phylo(hclust(daisy(cgMLSTdata_bird,
                                          metric="gower")))

@@

# Tree-stats
```{r, fig.width=10, fig.height=9}
print(paste0("Broiler median distance: ",
             round(median(cophenetic.phylo(cgMLST_tree_broiler)), 2)))

print(paste0("Pig median distance: ",
             round(median(cophenetic.phylo(cgMLST_tree_pig)), 2)))

print(paste0("Red fox median distance: ",
             round(median(cophenetic.phylo(cgMLST_tree_fox)), 2)))

print(paste0("Wild bird median distance: ",
             round(median(cophenetic.phylo(cgMLST_tree_bird)), 2)))

@@

hist(cophenetic.phylo(cgMLST_tree_broiler),
     breaks = 20,
     main = "Broiler distance frequency",
     xlab = "Distance",
     ylim = c(0,6000))

hist(cophenetic.phylo(cgMLST_tree_pig),
     breaks = 20,
     main = "Pig distance frequency",
     xlab = "Distance",
     ylim = c(0,6000))

hist(cophenetic.phylo(cgMLST_tree_fox),
     breaks = 20,
     main = "Red fox distance frequency",
     xlab = "Distance",
     ylim = c(0,6000))

hist(cophenetic.phylo(cgMLST_tree_bird),
     breaks = 20,
     main = "Wild bird distance frequency",
     xlab = "Distance",
     ylim = c(0,6000))

>>>>

## Specific mutations in non-QRDR intrinsic genes
These mutations are the most prominent in each gene where mutations were identified (n > 2). The percentage is based on how many of the total isolates with mutations in the respective gene had the specific mutation in question.
```{r, message=FALSE, warning=FALSE}
mut_filtered %>%
  filter(gene %not_in% c("gyrA","gyrB","parC","parE")) %>%
  select(id, gene, mut) %>%
  mutate(entries = sapply(strsplit(.$mut, ","),
                          FUN = function(x) {length(x)})) %>%
  separate(mut, into = paste("v", 1:max(.$entries), sep = "_"), sep = ",") %>%
  gather(mut, value, paste("v", 1:max(.$entries), sep = "_")) %>%
  mutate(value = str_trim(value)) %>%
  select(id, gene, value) %>%
  filter(value != 0,
         is.na(value) == FALSE) %>%
  group_by(gene, value) %>%
  dplyr::count() %>%
  ungroup() %>%
  filter(n > 2) %>%
  mutate(percent = case_when(
    gene == "gadX" ~ 
      round(n / length(which(mut_report[, "gadX"] == 1))*100, 1),
    gene == "marA" ~ 
      round(n / length(which(mut_report[, "marA"] == 1))*100, 1),
    gene == "marR" ~ 
      round(n / length(which(mut_report[, "marR"] == 1))*100, 1),
    gene == "rpoB" ~ 
      round(n / length(which(mut_report[, "rpoB"] == 1))*100, 1),
    gene == "soxS" ~ 
      round(n / length(which(mut_report[, "soxS"] == 1))*100, 1))) %>%
  arrange(gene, desc(percent)) %>%
  dplyr::rename("Gene" = gene,
                "Mutation" = value,
                "Percent" = percent)

@@

## Chi Squared test for significant differences in resistance mechanisms
```{r}
do_chisq_test <- function(df) {
  df <- df %>%
    dplyr::slice(expand.grid(1:length(unique(species)), 1:length(unique(species))) %>% rev %>% 
          dplyr::filter(Var2 < Var1) %>% t) %>%
    mutate(test = rep(1:(n()/2), each = 2)) %>%
    group_by(test) %>%
    dplyr::do(data_frame(gene = .$key,
                  test = dplyr::first(.$test),
                  species1 = dplyr::first(.$species),
                  species2 = dplyr::last(.$species),
                  data = list(matrix(c(.$Absent,
                                       .$Present),
                                     ncol = 2)))) %>%
    mutate(chi_test = map(data, chisq.test, correct = FALSE)) %>%
    mutate(p.value = map_dbl(chi_test, "p.value"),
           dupl = duplicated(test)) %>%
    dplyr::filter(dupl == FALSE) %>%
    ungroup() %>%
    select(gene, species1, species2, p.value) %>%
    mutate(p.value = round(p.value, 5)) %>%
    dplyr::filter(p.value <= 0.05)
  return(df)
}

@@

## Resistance mechanisms per sequence type
```{r}
mut_report %>%
  left_join(acquired_report, by = "id") %>%
  left_join(mlst_data[,c("id","ST")], by = "id") %>%
  select(-c(nt_change, id)) %>%
  gather(gene, value, -ST) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  filter(Present != 0) %>%
  mutate(Total = Present + Absent) %>%
  filter(Total > 2) %>%
  rowwise() %>%
  mutate(Percent = round(Present/Total*100, 1)) %>%
  ungroup() %>%
  group_by(ST) %>%
  mutate(no_of_mechanisms = n())
```

## Resistance mechanisms per sequence type per animal species
```{r}
mut_report %>%
  left_join(acquired_report, by = "id") %>%
  left_join(mlst_data[,c("id","ST")], by = "id") %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  select(-c(nt_change, id)) %>%
  gather(gene, value, -c(ST, species)) %>%
  group_by_all() %>%
  dplyr::count() %>%
  ungroup() %>%
  mutate(value = ifelse(value == 1, "Present", "Absent")) %>%
  spread(value, n, fill = 0) %>%
  filter(Present != 0) %>%
  mutate(Total = Present + Absent) %>%
  filter(Total > 2) %>%
  rowwise() %>%
  mutate(Percent = round(Present/Total*100, 1)) %>%
  ungroup() %>%
  group_by(ST, species) %>%
  mutate(no_of_mechanisms = n()) %>%
  arrange(ST, species)
```

@@

## Correlation between specific mutations and qnr-genes
```{r}
AddCol <- function(df, col_name) {
  # split rows by delimeters
  string_to_proc <- df %>% select(!!col_name) %>%
    base::unlist() %>% str_split(regex("\\, |\\,")) 
  # find unique entries
  unique_strings <- string_to_proc %>%
    base::unlist() %>% base::unique()
  # construct names of the new columns
  cols_names <- paste(col_name, unique_strings, sep = "_")
  # construct 0/1-content columns for each unique entry
  cols_content <- sapply(function(i) {
    as.integer(base::unlist(lapply(function(Z) any(Z %in% unique_strings[i]), 
                             X = string_to_proc)))
  }, X = seq_along(unique_strings))
  res <- data.frame(cols_content)
  names(res) <- cols_names
  return(res)
}

create_corr_matrix <- function(df) {
  df <- df %>%
    as.matrix %>%
    cor(method = "spearman") %>%
    as.data.frame %>%
    rownames_to_column(var = "var1") %>%
    gather(var2, value, -var1) %>%
    mutate(var1 = gsub("(.*?)_R$", "\\1", var1),
           var2 = gsub("(.*?)_R$", "\\1", var2))
  return(df)
}

col_proc <- c("gyrA","gyrB","parC","parE")

cols_to_add <- sapply(function(i) {AddCol(df = mut_quant, col_name = col_proc[i])},
                      X = seq_along(col_proc)) %>%
  bind_cols() %>%
  select(-c(gyrA_0, parC_0, gyrB_0, parE_0)) %>%
  mutate_all(funs(as.numeric)) %>%
  {. ->> tot_mut_report} %>%
  select(gyrA_S83L, gyrA_D87N, parC_S80I) %>%
  mutate(gyrA_S83L_D87N = ifelse(gyrA_S83L == 1 & gyrA_D87N == 1, 1, 0),
         gyrA2_parC1 = ifelse(gyrA_S83L == 1 & gyrA_D87N == 1 & parC_S80I == 1, 1, 0))

test <- mut_quant %>%
  cbind(tot_mut_report) %>%
  select(-contains("mut")) %>%
  select(-c(gyrA, gyrB, parC, parE)) %>%
  left_join(acquired_report, by = "id") %>%
  mutate_at(vars(-id),
            funs(as.numeric)) %>%
  mutate(qnr = ifelse(rowSums(.[,21:27]) == 0, 0, 1)) %>%
  select(-id)

test2 <- create_corr_matrix(test)

cor.test(test$gyrA_S83L, test$qnr)
```

@@

## Correlation plot for resistance mechanisms
```{r, fig.height=8, fig.width=7, warning=FALSE}
total_df <- acquired_report %>%
  left_join(mut_report, by = "id") %>%
  left_join(isolate_data[c("id","species")], by = "id") %>%
  select(-c(nt_change, id)) %>%
  mutate_at(vars(-species),
            funs(as.numeric)) %>%
  mutate(qnr = ifelse(rowSums(.[,2:8]) > 0, 1, 0)) %>%
  {. ->> total_resistance_report} %>%
  split(., f = .$species) %>%
  lapply(., function(x) select(x, -species))

corr_matrices <- lapply(total_df, function(x) create_corr_matrix(x)) %>%
  bind_rows(.id = "species") %>%
  mutate(gene1 = gsub("(.*?)_.+", "\\1", var1),
           gene2 = gsub("(.*?)_.+", "\\1", var2),
           var1 = gsub(".+_(.*?)", "\\1", var1),
           var2 = gsub(".+_(.*?)", "\\1", var2),
           type1 = case_when(gene1 == "gyrA" ~ 1,
                          gene1 == "gyrB" ~ 2,
                          gene1 == "parC" ~ 3,
                          gene1 == "parE" ~ 4,
                          gene1 == "marR" ~ 5,
                          gene1 == "marA" ~ 6,
                          gene1 == "rpoB" ~ 7,
                          gene1 == "soxS" ~ 8,
                          gene1 == "robA" ~ 9,
                          gene1 == "gadX" ~ 10,
                          gene1 == "qnrA1" ~ 11,
                          gene1 == "qnrB19" ~ 12,
                          gene1 == "qnrB6" ~ 13,
                          gene1 == "qnrB60" ~ 14,
                          gene1 == "qnrS1" ~ 15,
                          gene1 == "qnrS2" ~ 16,
                          gene1 == "qnrS4" ~ 17,
                          gene1 == "qnr" ~ 18,
                          gene1 == "qepA4" ~ 19),
           type2 = case_when(gene2 == "gyrA" ~ 1,
                          gene2 == "gyrB" ~ 2,
                          gene2 == "parC" ~ 3,
                          gene2 == "parE" ~ 4,
                          gene2 == "marR" ~ 5,
                          gene2 == "marA" ~ 6,
                          gene2 == "rpoB" ~ 7,
                          gene2 == "soxS" ~ 8,
                          gene2 == "robA" ~ 9,
                          gene2 == "gadX" ~ 10,
                          gene2 == "qnrA1" ~ 11,
                          gene2 == "qnrB19" ~ 12,
                          gene2 == "qnrB6" ~ 13,
                          gene2 == "qnrB60" ~ 14,
                          gene2 == "qnrS1" ~ 15,
                          gene2 == "qnrS2" ~ 16,
                          gene2 == "qnrS4" ~ 17,
                          gene2 == "qnr" ~ 18,
                          gene2 == "qepA4" ~ 19))

@@

ggplot(corr_matrices, aes(reorder(var1, type1), reorder(var2, type2), fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient(low = "black",
                      high = "white",
                      na.value = "grey95")+
  labs(fill = NULL)+
  theme_classic()+
  theme(axis.text = element_text(size = 8),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.4),
        axis.title = element_blank())+
  coord_fixed()+
  facet_wrap(~species)
```

@@

# Dendrograms

## Number of NA's per isolate in cgMLST allele matrix
The red outliers have been excluded from the cgMLST analysis due to too many missing alleles
```{r}
cgMLSTdata <- read.table("../data/cgMLST.txt",
                         sep = "\t",
                         colClasses = "factor",
                         header = T)

cgMLSTdata[cgMLSTdata == "NaN"] <- NA

cgMLSTdata2 <- cgMLSTdata %>%
  mutate(n_NA = apply(., 1, function(x) sum(is.na(x))),
         group = 1)

ggplot(cgMLSTdata2, aes(factor(group), n_NA)) +
  geom_boxplot(outlier.size = 2) +
  geom_point(data = cgMLSTdata2[cgMLSTdata2$n_NA > 500,], aes(factor(group), n_NA),
             pch = 21,
             size = 2,
             fill = "red") +
  labs(x = NULL,
       y = "Number of NA's per isolate") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

## cgMLST-tree total
```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
palette3 <- c("1-A" = "#fc8d59",
              "1-I" = "#80b1d3",
              "0" = "grey95",
              "0.03" = "#440154FF",
              "0.06" = "#482878FF",
              "0.12" = "#3E4A89FF",
              "0.25" = "#31688EFF",
              "0.5" = "#26828EFF",
              "1" = "#1F9E89FF",
              "2" = "#35B779FF",
              "4" = "#6DCD59FF",
              "8" = "#B4DE2CFF",
              "16" = "#FDE725FF")

total_data_id <- total_data %>%
  column_to_rownames("id")

metadata_total <- mlst_data[,c("id","ST")] %>%
  left_join(isolate_data[,c("id", "species")], by = "id")

cgMLSTdata_complete <- cgMLSTdata2 %>%
  filter(n_NA < 500) %>%
  select(-n_NA) %>%
  column_to_rownames("Genome")

total_tree <- as.phylo(hclust(daisy(cgMLSTdata_complete, metric = "gower"), method = "average"))

total_tree_annotated <- ggtree(total_tree,
                               layout = "circular",
                               color = "#808080",
                               size = 0.5) %<+% metadata_total +
                               geom_tippoint(aes(color = species),
                               size = 3) +
                               geom_tiplab2(aes(label = ST),
                               offset = 0.01,
                               size = 4) +
  scale_color_manual(values = palette)

total_tree_opened <- rotate_tree(open_tree(total_tree_annotated, 8), 93)

gheatmap(total_tree_opened,
         total_data_id,
         offset = 0.04,
         width = 0.3,
         colnames_offset_y = 2.5,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette3,
                    labels = c("Gene/Mutation absent",
                               "Acquired gene present",
                               "Intrinsic gene mutated")) +
  theme(legend.position = c(0.9,0.9),
        legend.text = element_text(size = 20))

## Roary tree
```{r, fig.height=20, fig.width=20, message=FALSE, warning=FALSE}
roary_tree <- read.tree("../data/Roary/accessory_binary_genes.fa.newick")

tree_refs <- roary_tree$tip.label

id_data2 <- id_data %>%
  mutate(test = ref %in% tree_refs) %>%
  filter(test == TRUE)

id_df <- left_join(data.frame(ref = tree_refs), id_data2, by = "ref") %>%
  mutate(id = ifelse(is.na(id) == TRUE, ref, id))

roary_tree$tip.label <- id_df$id

total_data2 <- total_data %>%
  filter(id %in% id_df$id) %>%
  column_to_rownames("id")

metadata_total2 <- metadata_total %>%
  filter(id %in% id_df$id)

roary_tree_colored <- ggtree(roary_tree,
       layout = "circular",
       color = "#808080",
       size = 1) %<+% metadata_total2 +
  geom_tiplab2(aes(label = ST, color = species),
               offset = 0.01,
               size = 6,
               align = TRUE) +
  scale_color_manual(values = palette)

total_tree_opened2 <- rotate_tree(open_tree(roary_tree_colored, 8), 93)

gheatmap(total_tree_opened2,
         total_data2,
         offset = 0.065,
         width = 0.3,
         colnames_offset_y = 1.9,
         colnames_position = "top",
         font.size = 5) +
  scale_fill_manual(values = palette3,
                    labels = c("Gene/Mutation absent",
                               "Acquired gene present",
                               "Intrinsic gene mutated")) +
  theme(legend.position = c(0.9,0.9),
        legend.text = element_text(size = 20))

## cgMLST-tree per animal species
```{r, fig.width=28, fig.height=28, message = FALSE, warning = FALSE}

@@

cgMLSTdata <- read.table("../data/cgMLST.txt",
                         sep = "\t",
                         colClasses = "factor",
                         header = T)

cgMLSTdata[cgMLSTdata == "NaN"] <- NA

cgMLSTdata2 <- cgMLSTdata %>%
  mutate(n_NA = apply(., 1, function(x) sum(is.na(x)))) %>%
  filter(n_NA < 500) %>%
  select(-n_NA) %>%
  column_to_rownames("Genome")

@@

rawdata_broiler <- cgMLSTdata2 %>%
  rownames_to_column("id") %>%
  filter(id %in% broiler_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
  {. ->> data_broiler } %>%
  select(-ST)

rawdata_pig <- cgMLSTdata2 %>%
  rownames_to_column("id") %>%
  filter(id %in% pig_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
  {. ->> data_pig } %>%
  select(-ST)

rawdata_fox <- cgMLSTdata2 %>%
  rownames_to_column("id") %>%
  filter(id %in% fox_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
  {. ->> data_fox } %>%
  select(-ST)

rawdata_bird <- cgMLSTdata2 %>%
  rownames_to_column("id") %>%
  filter(id %in% bird_iso$id) %>%
  dplyr::rename("Genome" = `id`) %>%
  select(Genome, everything()) %>%
  left_join(mlst_data[c("id","ST")], by = c("Genome" = "id")) %>%
  {. ->> data_bird } %>%
  select(-ST)

write.table(rawdata_broiler, "../data/cgMLST_broiler.txt", sep = "\t", row.names = FALSE)
write.table(rawdata_pig, "../data/cgMLST_pig.txt", sep = "\t", row.names = FALSE)
write.table(rawdata_fox, "../data/cgMLST_fox.txt", sep = "\t", row.names = FALSE)
write.table(rawdata_bird, "../data/cgMLST_bird.txt", sep = "\t", row.names = FALSE)

@@

cgMLSTdata_broiler[cgMLSTdata_broiler == "NaN"] <- NA

cgMLST_tree_broiler <- as.phylo(hclust(daisy(cgMLSTdata_broiler,
                                             metric="gower"), method = "average"))

@@

cgMLSTdata_pig[cgMLSTdata_pig == "NaN"] <- NA

cgMLST_tree_pig <- as.phylo(hclust(daisy(cgMLSTdata_pig,
                                         metric="gower"), method = "average"))

@@

cgMLSTdata_fox[cgMLSTdata_fox == "NaN"] <- NA

cgMLST_tree_fox <- as.phylo(hclust(daisy(cgMLSTdata_fox,
                                         metric="gower"), method = "average"))

@@

cgMLSTdata_bird[cgMLSTdata_bird == "NaN"] <- NA

cgMLST_tree_bird <- as.phylo(hclust(daisy(cgMLSTdata_bird,
                                          metric="gower"), method = "average"))

@@

# Tree-stats
```{r, fig.width=10, fig.height=9}
dist_broiler <- cophenetic.phylo(cgMLST_tree_broiler)
dist_broiler[dist_broiler > 0.95] <- NA

dist_pig <- cophenetic.phylo(cgMLST_tree_pig)
dist_pig[dist_pig > 0.95] <- NA

dist_fox <- cophenetic.phylo(cgMLST_tree_fox)
dist_fox[dist_fox > 0.95] <- NA

dist_bird <- cophenetic.phylo(cgMLST_tree_bird)
dist_bird[dist_bird > 0.95] <- NA

print(paste0("Broiler median distance: ",
             round(median(dist_broiler, na.rm = TRUE), 2)))

print(paste0("Pig median distance: ",
             round(median(dist_pig, na.rm = TRUE), 2)))

print(paste0("Red fox median distance: ",
             round(median(dist_fox, na.rm = TRUE), 2)))

print(paste0("Wild bird median distance: ",
             round(median(dist_bird, na.rm = TRUE), 2)))

@@

x1 <- hist(dist_broiler,
     breaks = 20,
     plot = FALSE)

x1$density = x1$counts/sum(x1$counts)*100

plot(x1,
     freq = FALSE,
     main = "Broiler distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

x2 <- hist(dist_pig,
     breaks = 20,
     plot = FALSE)

x2$density = x2$counts/sum(x2$counts)*100

plot(x2,
     freq = FALSE,
     main = "Pig distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

x3 <- hist(dist_fox,
     breaks = 20,
     plot = FALSE)

x3$density = x3$counts/sum(x3$counts)*100

plot(x3,
     freq = FALSE,
     main = "Red fox distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))

x4 <- hist(dist_bird,
     breaks = 20,
     plot = FALSE)

x4$density = x4$counts/sum(x4$counts)*100

plot(x4,
     freq = FALSE,
     main = "Wild bird distance frequency",
     xlab = "Distance",
     ylim = c(0, 50))
